---
title: "Dissertation "
author: "Samid Ali"
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 2
    
---

```{r setup, include=FALSE}



library(ggplot2)
library(knitr)
library(ivreg)
library(ggdag)
library(data.table)
library(dplyr)
library(tidyr)
library(stargazer)
library(clipr)
library(tibble)
library(lubridate)
# library(kableExtra) 

# knitr::opts_chunk$set(echo = TRUE)


```


\newpage


# 1  

VAR

We estimate the following model by OLS:     
$$
\begin{aligned}
loggdp_{i} = \beta_0 + \beta_1 Risk + X'\alpha + u_i
\end{aligned}
$$
Where:   
$Risk$ is the measure of protection against expropriation,  
$X$ is a vector of controls (in the simple linear regression model this is assumed to be the 0 vector, while in the multiple linear regression model we include edesc1975)  
$u_i$ is the unobserved error term,

<!-- sessionInfo() -->


```{r 1, echo=FALSE}

df <- fread("D:/Samid work/University/KCL - Econ and Policy/Dissertation/Data/GDP.csv", skip = 1)


# Filter the data frame to exclude rows where the column 'title' matches any of the specified values
filtered_df <- df %>% 
  # Only keep the quarterly data
  subset(nchar(CDID) == 7 & substr(CDID,6,6) == "Q") %>% 
  select(CDID,Deflator = L8GG, GDP = ABMI) %>% 
  mutate(Year = as.numeric(substr(CDID,1,4)),
         Quarter = substr(CDID,6,7),
         Q  = as.numeric(substr(CDID,7,7))) %>% 
  # For testing purposes, can change later
  subset(Year >= 1987)  %>% 
  mutate(Deflator = as.numeric(Deflator),
         GDP = as.numeric(GDP))


# View the filtered data frame
# head(filtered_df)

str(filtered_df)


# L8GG

```


```{r dataProc, echo=FALSE}

fiscal_raw <- fread("D:/Samid work/University/KCL - Econ and Policy/Dissertation/Data/Fiscal Data.csv", skip = 1)

fiscal_proc <- fiscal_raw %>% 
  select(Date_ID = Transaction, Revenue = OTR, Expenditure = OTE) %>% 
  subset(Date_ID !=  "Dataset identifier code" & Date_ID != "Identifier") %>% 
  mutate(Year = as.numeric(gsub("\\D", "", Date_ID)),
         Period = gsub("\\d{4}", "", Date_ID))  %>% 
  mutate(
    Q = case_when(
      Period == "Jan to Mar " ~ 1,
      Period == "Apr to Jun " ~ 2,
      Period == "Jul to Sep " ~ 3,
      Period == "Oct to Dec " ~ 4
      ),
    Unique_Period = Year +(Q/4)
   
  ) %>% 
  # Convert to numeric and multiply by 1 million so values as these will later be made into per capita terms
  mutate(Revenue = as.numeric(gsub(",", "", Revenue) ),
         Expenditure = as.numeric(gsub(",", "", Expenditure )))

# join GDP deflator and GDP data

population <- fread("D:/Samid work/University/KCL - Econ and Policy/Dissertation/Data/Population.csv", 
                    skip = 4, 
                    header = TRUE) %>% 
  subset(`Country Name` == "United Kingdom") %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Year") %>% 
  rename(Population = V1 ) %>% 
  filter(grepl("^\\d{4}$", Year)) %>% 
  mutate(Year = as.numeric(Year),
         Population = as.numeric(Population))

Interest_SR <- fread("D:/Samid work/University/KCL - Econ and Policy/Dissertation/Data/3 Month Treasurey Rate.csv") %>% 
  mutate(Date = dmy(Date),
         month = month(Date),
         Year = year(Date),
         Q = case_when(
           month %in% 1:3 ~ 1,
           month %in% 4:6 ~ 2,
           month %in% 7:9 ~ 3,
           month %in% 10:12 ~ 4
         )) %>% 
  group_by(Year, Q) %>% 
  summarize(mean_SR_Rate = mean(SR_Rate, na.rm = TRUE))


data <- fiscal_proc %>% 
  left_join(filtered_df, by = c("Q" = "Q", "Year" = "Year")) %>% 
  left_join(Interest_SR, by = c("Q" = "Q", "Year" = "Year")) %>% 
  left_join(population, by = c("Year" = "Year")) %>% 
# Convert variables to per capita, note revenue, expenditure, and GDP are in £ million so need to multiply. Doing this
  mutate(RevenuePerCapita = (Revenue *10^6) /Population,
         ExpenditurePerCapita = (Expenditure *10^6) /Population,
         GDPPerCapita = (GDP *10^6) /Population) %>% 
# Convert variables (except interest rate) to logs
  mutate(log_revenue = log(Revenue *10^6),
         log_expenditure = log(Expenditure *10^6),
         log_GDP = log(GDP *10^6),
         log_deflator = log(Deflator))

         
model_data <- data %>% 
  select(CDID, log_revenue, log_expenditure, log_GDP, log_deflator, mean_SR_Rate)




```

```{r , echo=FALSE}

# ggplot(fiscal_proc, aes(x = Year2)) +
#   geom_line(aes(y = Revenue, color = "Revenue"), size = 1) +
#   geom_line(aes(y = Expenditure, color = "Expenditure"), size = 1) +
#   labs(
#     x = "Date ID",
#     y = "Amount (? in millions)",
#     title = "Revenue and Expenditure Over Time",
#     color = "Legend"
#   ) +
#   scale_color_manual(values = c("Revenue" = "blue", "Expenditure" = "red")) +
#   theme_minimal(base_size = 15) +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     plot.title = element_text(hjust = 0.5, face = "bold"),
#     axis.title.x = element_text(face = "bold"),
#     axis.title.y = element_text(face = "bold"),
#     legend.position = "bottom",
#     legend.title = element_text(face = "bold")
#   )


```



Identification:

# Recovering Structural Shocks

Let \( Y_t \) be the vector of variables:

\[ Y_t = \begin{pmatrix} G_t \\ R_t \\ GDP_t \\ T_t \\ P_t \end{pmatrix} \]

The reduced-form VAR model can be written as:

\[ Y_t = A_1 Y_{t-1} + A_2 Y_{t-2} + \cdots + A_p Y_{t-p} + \epsilon_t \]

where \( \epsilon_t \) is the vector of reduced-form residuals. To recover the structural shocks \( u_t \), we assume:

\[ \epsilon_t = B u_t \]

where \( B \) is a lower triangular matrix. The structural shocks \( u_t \) are assumed to be uncorrelated and have unit variance.

The matrix \( B \) can be obtained using Cholesky decomposition of the covariance matrix of the reduced-form residuals \( \Sigma_\epsilon \):

\[ \Sigma_\epsilon = E[\epsilon_t \epsilon_t'] = BB' \]

Given the recursive ordering \( (G, R, GDP, T, P) \), the matrix \( B \) has the form:

\[ B = \begin{pmatrix} 
b_{11} & 0 & 0 & 0 & 0 \\ 
b_{21} & b_{22} & 0 & 0 & 0 \\ 
b_{31} & b_{32} & b_{33} & 0 & 0 \\ 
b_{41} & b_{42} & b_{43} & b_{44} & 0 \\ 
b_{51} & b_{52} & b_{53} & b_{54} & b_{55} 
\end{pmatrix} \]

Thus, the structural shocks \( u_t \) can be recovered as:

\[ u_t = B^{-1} \epsilon_t \]

